Spanner: Google's Globally-Distributed Database \newline Seminar Report

### 2016-02-16 paper/paper.tex
Iterate TrueTime

\section{TrueTime \tbr}
\label{sec:truetime}

This section describes the key enabler of Spanner, i.e. the novel \emph{TrueTime}
API. Essentially, this API enables Google to provide timestamps for transactions
which reflect serialization order.

TrueTime represents time as a $TTinterval$. A $TTinterval$ is an interval with a
bounded time uncertainty and has two endpoints, i.e. $earliest$ and $latest$,
which are of type $TTstamp$ (which is a timestamp in the context of Spanner).
Table~\ref{tbl:truetime-api} summarizes the three methods provided by the TrueTime
API. However, the most important method is $TT.now()$ which returns a $TTinterval$
which is guaranteed to contain the absolute time of the invocation of $TT.now()$,
i.e. $TT.now().earliest \leq t_{abs}(e) \leq TT.now().latest$, where $e$ denotes
the event of invoking $TT.now()$. More precisely, the absolute time of invocation
of $TT.now()$ is exactly in the middle this $TTinterval$ and spans over
$2\cdot\epsilon$ where $\epsilon$ denotes the instantaneous error bound. $\epsilon$
is derived from applied worst-case local clock drift (assumed to be
$200\mu$s/sec). This instantaneous error bound is adjusted over time for every
single machine.

TrueTime uses GPS and atomic clocks as time references. Google chose these two
types of time reference because the failure modes of them are uncorrelated. A
datacenter typically consists of multiple \emph{time master} machines. Most of
these are GPS receivers with dedicated antennas. All remaining time master
machines use an atomic clock as reference. The references of all time masters are
compared against each other on a regular basis. To avoid incorrect time references
time masters also cross-check their local clock drift against the clock drift of
the reference. If these differ to much, the time master evicts itself.

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{|l||l|}
    \hline
    {\bfseries Method} & {\bfseries Returns} \tabularnewline
    \hline\hline
    $TT.now()$ & $TTinterval$: [$earliest$, $latest$] \tabularnewline
    \hline
    $TT.after(t)$ & true if $t$ has definitely passed \tabularnewline
    \hline
    $TT.before(t)$ & true if $t$ has definitely now arrived \tabularnewline
    \hline
  \end{tabular}
  \caption{Methods provided by the TrueTime API}
  \label{tbl:truetime-api}
\end{table}

@@

The system introduced by Google is definitly impressive because they were able
to establish a notion of global wall clock time (that's how they call it in their
presentation) which enabled them to guarantee serialization order for
globally-distributed transactions. To the best of the author's knowledge, Spanner
was the first system to provide much stronger consistency and time semantics while
data is sharded and replicated at global scale. Google also proved all
'\emph{Relational databases are dead}' people wrong even though systems like
Bigtable helped creating the NoSQL trend. Reading about a system like Spanner is
definitly more than a vital sign of relational database systems. From a
developer's point of view, this is definitly interesting because Spanner provides
the more intuitive schematized view of tables. In contrast to NOSQL database
systems, Spanner could be used as replacement for other relational database
systems, e.g. MySQL, directly. Spanner is also highly flexible to schema changes
because of the atomic schema change transactions.

>>>>

\section{TrueTime}
\label{sec:truetime}

This section describes the key enabler of Spanner, i.e. the novel \emph{TrueTime}
API. Essentially, this API enables Google to assign timestamps to distributed
transactions which reflect serialization order.

TrueTime represents time as a $TTinterval$. A $TTinterval$ is an interval with a
bounded time uncertainty and has two endpoints, i.e. $earliest$ and $latest$,
which are of type $TTstamp$ (which is a timestamp in the context of Spanner).
Table~\ref{tbl:truetime-api} summarizes the three methods provided by the TrueTime
API. However, the most important method is $TT.now()$ which returns a $TTinterval$
which is guaranteed to contain the absolute time of the invocation of $TT.now()$,
i.e. $TT.now().earliest \leq t_{abs}(e) \leq TT.now().latest$, where $e$ denotes
the event of invoking $TT.now()$. More precisely, the absolute time of invocation
of $TT.now()$ is exactly in the middle this $TTinterval$. The interval itself
spans over $2\cdot\epsilon$, where $\epsilon$ denotes the instantaneous error
bound. $\epsilon$ is derived from applied worst-case local clock drift (assumed
to be $200\mu$s/sec). This instantaneous error bound is adjusted over time for
every single machine.

TrueTime uses GPS and atomic clocks as time references. Google chose these two
types of time references because the failure modes of them are uncorrelated. A
datacenter typically consists of multiple \emph{time master} machines. Most of
these are GPS receivers with dedicated antennas. All remaining time master
machines use an atomic clock as reference. Machines using an atomic clock as time
reference are referred to as \emph{Armageddon} masters. The references of all time
masters are compared against each other on a regular basis. To avoid incorrect
time references, time masters also cross-check their local clock drift against
the clock drift of the reference. If these differ too much, the time master evicts
itself.

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{|l||l|}
    \hline
    {\bfseries Method} & {\bfseries Returns} \tabularnewline
    \hline\hline
    $TT.now()$ & $TTinterval$: [$earliest$, $latest$] \tabularnewline
    \hline
    $TT.after(t)$ & true if $t$ has definitely passed \tabularnewline
    \hline
    $TT.before(t)$ & true if $t$ has definitely not arrived \tabularnewline
    \hline
  \end{tabular}
  \caption{Methods provided by the TrueTime API.}
  \label{tbl:truetime-api}
\end{table}

@@

The system introduced by Google is definitely impressive because they were able
to establish a notion of global wall clock time (that's how they call it in their
presentation) which enabled them to guarantee serialization order for
globally-distributed transactions. To the best of the author's knowledge, Spanner
was the first system to provide much stronger consistency and time semantics while
data is sharded and replicated at global scale. Google also proved all
'\emph{Relational databases are dead}' people wrong even though systems like
Bigtable helped creating the NoSQL trend. Reading about a system like Spanner is
definitely more than a vital sign of relational database systems. From a
developer's point of view, this is interesting because Spanner provides
the more intuitive schematized view of tables. In contrast to NOSQL database
systems, Spanner could be used as replacement for other relational database
systems, e.g. MySQL, directly. Spanner is also highly flexible to schema changes
because of the atomic schema change transactions.
